---
title: "NIRS_rev1"
author: "Slaven Pikija"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

## Results

Among `r n_total` patients, rSO₂ was lower on the affected side in
`r n_unaff_higher` (`r sprintf("%.1f", pct_unaff_higher)`%),
similar between hemispheres in `r n_equal`
(`r sprintf("%.1f", pct_equal)`%), and higher on the affected side in
`r n_aff_higher` (`r sprintf("%.1f", pct_aff_higher)`%).

The distribution across the three categories differed significantly from an equal distribution
(χ² = `r sprintf("%.2f", unname(chi_test$statistic))`,
df = `r unname(chi_test$parameter)`,
p < `r format.pval(chi_test$p.value, digits = 3, eps = 0.001)`).

Among the `r n_determinate` patients with a directional interhemispheric
difference, the hemisphere with lower rSO₂ correctly corresponded to the
angiographically affected hemisphere in `r n_unaff_higher`
(`r sprintf("%.1f", ident_accuracy)`%;
95% CI `r sprintf("%.1f", ident_ci_low)`–`r sprintf("%.1f", ident_ci_high)`%;
exact binomial p = `r format.pval(binom_test$p.value, digits = 3, eps = 0.001)`).

## Figure

```{r affected-hemisphere-dotplot, echo=FALSE, fig.width=5, fig.height=8, dpi=300}
fig1
```

Supplemental Figure xx.
Interhemispheric difference in mean regional cerebral oxygen saturation (rSO₂) during the prespecified 10-minute analysis period. ΔrSO₂ was calculated as affected minus unaffected hemisphere. The shaded area indicates the prespecified range considered similar between hemispheres (−4 to +4 percentage points). Positive values indicate higher rSO₂ on the affected side, and negative values indicate higher rSO₂ on the unaffected side. N=170.